---
title: "Week 8 Wednesday Worksheet"
author: "Valentino Aceves"
date: "May 22, 2024"
output: pdf_document
---

# Joint Hypothesis Testing

$$
Y =\beta_0 + \beta_1X_1 + \beta_2X_2 + \beta_3X_3 + \beta_4X_4
$$
An example of a Joint Null Hypothesis could be
$$
H_0: \beta_1 =\beta_2=0\ \ \ \ \ , \ \ \ \ H_1:\text{at least one of the estimates}\ne0
$$
We can now create two different models:

**Unrestricted Regression** which is our original:
$$
Y =\beta_0 + \beta_1X_1 + \beta_2X_2 + \beta_3X_3 + \beta_4X_4
$$

**Restricted Regression** where we assume $H_0$ is true:
$$
Y =\beta_0 + \beta_3X_3 + \beta_4X_4
$$

Let $SSR_u$, $SSR_r$ be the $SSR$'s of the unrestricted and restricted models respectively. Let $p=4$ be the number of parameters in the unrestricted model. Let $q=2$ be the number of parameters in the restricted model. 
We calculate our $F$ statistic with
$$
\hat{F} = \frac{(SSR_r - SSR_u)/(p-q)}{SSR_u/(n-p-1)} \sim F_{df_1,df_2}
$$
where $df_1 = p-q$ and $df_2 = n-p-1$. We reject $H_0$ when $\hat{F} > F_{df_1,df_2}^c$ or $p_{\hat{F}} < \alpha$.

\newpage

# Example 1

```{r, message=FALSE}
library(POE5Rdata)
data("cocaine")
```
```{r}
names(cocaine)
```


```{r}
unrestricted_model <- lm(price ~ quant + qual + trend, cocaine)
restricted_model <- lm(price ~ quant, cocaine)
```
### Perform an F test with the restricted and unrestricted models from above. State $H_0$ and $H_1$.

```{r}
df1 <- 3 - 1 # p-q
df2 <- unrestricted_model$df.residual # n-p-1
c(df1, df2)
```

```{r}
SSR_u <- sum(unrestricted_model$residuals^2)
SSR_r <- sum(restricted_model$residuals^2)
c(SSR_u, SSR_r)
```

```{r}
Fc <- qf(0.95, df1, df2)
F_stat <- ((SSR_r - SSR_u)/df1) / (SSR_u/df2)
c(Fc, F_stat)
```


```{r}
1 - pf(F_stat, df1, df2)
```

\newpage

# F-tests on Single Variables

Consider these two model for an F-test:

**Unrestricted Regression**
$$
Y =\beta_0 + \beta_1X_1 + \beta_2X_2 + \beta_3X_3 
$$

**Restricted Regression**
$$
Y =\beta_0 + \beta_1X_1 + \beta_2X_2
$$



```{r}
unrestricted_model <- lm(price ~ quant + qual + trend, cocaine)
restricted_model <- lm(price ~ quant + qual, cocaine)
```
### Perform an F test with the restricted and unrestricted models from above. State $H_0$ and $H_1$.

```{r}
df1 <- 3 - 2 # p-q
df2 <- unrestricted_model$df.residual # n-p-1
SSR_u <- sum(unrestricted_model$residuals^2)
SSR_r <- sum(restricted_model$residuals^2)
```

```{r}
Fc <- qf(0.95, df1, df2)
F_stat <- ((SSR_r - SSR_u)/df1) / (SSR_u/df2)
c(Fc, F_stat)
```

Couldn't we also do a t-test on the single variable? Sure.

```{r}
summary(unrestricted_model)
```

We can see the t-statistic for $H_0: \beta_3 = 0, H_1:\beta_3\ne0$ is $\hat{t}=-1.699$. 

```{r}
summary(unrestricted_model)$coefficients[4, 3] ^ 2 # t_hat squared
```

In a single variable hypothesis test, when we square the t-statistic, we get exactly the F-statistic.

\newpage

# F-Tests for Overall Significance

When we run a summary table, it gives us an F-stat along with its p-value. What is the F-test that the regression is running for us?

$$
H_0:\beta_1=\beta_2=\cdots=\beta_k=0 \ \ \ \ \ \ \ \text{vs}\ \ \ \ \ \ \ H_1:\text{at least one } \beta_i\ne0
$$
Which leaves us with a restricted model of
$$
Y=\beta_0\ \ \ \ \ \ \implies\ \ \ \ \ \beta_0=\bar{Y}
$$
$$
\implies SSR_r = \sum_{i=1}^n (y_i-\hat{y}_i)^2 = \sum_{i=1}^n (y_i-\bar{y})^2 = TSS
$$

### Prove that, in this model, $F$ statistic becomes

$$
\hat{F}=\frac{R^2/p}{(1-R^2)/(n-p-1)}
$$


