---
title: "Week 6 Monday Worksheet"
author: "Valentino Aceves"
date: "May 6th, 2024"
output: pdf_document
---

```{r}
library(POE5Rdata)
```

```{r}
data(wa_wheat)
names(wa_wheat)
```
```{r}
model <- lm(chapman ~ time, data = wa_wheat)
resid <- model$residuals
plot(resid ~ wa_wheat$time, col = "pink", pch = 18)
```

```{r}
library(tseries)
```
```{r}
jarque.bera.test(resid)
```
We fail to reject the null hypothesis. The residuals are distributed normally "enough".

```{r part 5b}
model_linlog <- lm(chapman ~ log(time), data = wa_wheat)
resid_linlog <- model_linlog$residuals
plot(resid_linlog ~ log(wa_wheat$time), pch = 18, col = "pink")
```

```{r}
jarque.bera.test(resid_linlog)
```

## Part C

```{r}
summary(model)
summary(model_linlog)
```

When comparing R2 the linear model explains more of the variance in chapman than the linear log model. When we look at the jbtest, the linlog model has a lower pvalue. this means the errors are distributed less normally than the errors in the linear model. talk about plots. plot the fitted equations against the data. You guys make the claims, here is the evidence.

<!-- plotting fitted line -->

```{r plot linear}
plot(wa_wheat$chapman ~ wa_wheat$time)
abline(model)
```
```{r plot linlog}
plot(wa_wheat$chapman ~ wa_wheat$time)
curve(model_linlog$coefficients[1] + model_linlog$coefficients[2] * log(x), 
      add = TRUE, lwd = 2, col = "skyblue")
```




# Question 4

```{r part 4a}
m1_b0 <- 3.446
m1_b1 <- -0.001459
m1_yhat <- m1_b0 + m1_b1 * (10 - 35)^2
m1_yhat
```

```{r}
m2_b0 <- 1.4276
m2_b1 <- 0.5343
m2_yhat <- m2_b0 + m2_b1 * log(10)
m2_yhat
```

```{r part b}
m1_marg <- 2 * m1_b1 * (10 - 35)
m1_marg
m2_marg <- m2_b1 / 10
m2_marg
```

### part c

$$
\lambda_1 = 2\cdot \beta_1 \cdot (EXPER - 35)\\
SE(\lambda_1) = \sqrt{Var(\lambda_1)} = \sqrt{Var(2\cdot \beta_1 \cdot (EXPER - 35))}
$$
```{r}
se_m1 <- sqrt(2^2 * (10-35)^2 * (0.0000786)^2)
se_m1
tc <- qt(0.975, 48)
m1_yhat + c(-tc, tc) * se_m1
```

```{r}
se_m2 <- sqrt((1/10)^2 * (0.04333)^2)
se_m2
m2_yhat + c(-tc, tc) * se_m2
```




